---
title: "RLT Package Testing Survival Functions and Features"
author: "Ruoqing Zhu"
date: "Last Updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true 
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r set-options, echo=FALSE, cache=FALSE}
  options(width = 1000)
  knitr::opts_chunk$set(fig.width=9, fig.height=7, out.width = "75%", fig.align = 'center')
  knitr::opts_chunk$set(class.source = "fold-show")
  knitr::opts_chunk$set(collapse=TRUE)
```

## Install and Load Package

Install and load the GitHub version of the RLT package. Do not use the CRAN version. 

```{r}
  # install.packages("devtools")
  # devtools::install_github("teazrq/RLT")
  library(RLT)
```

## Confidence Band Estimation

```{r}
  # generate a simpler data
  library(RLT)
  set.seed(2)
  n = 600
  p = 20
  X = matrix(rnorm(n*p), n, p)
  
  xlink <- function(x) exp(x[, 1] + x[, 3]/2)
  FT = rexp(n, rate = xlink(X))
  CT = pmin(6, rexp(n, rate = 0.25))
  
  Y = pmin(FT, CT)
  Censor = as.numeric(FT <= CT)
  mean(Censor)
  
  ntest = 15
  testx = matrix(rnorm(ntest*p), ntest, p)
  #testx[1, 1:3] = c(0.5, 0, 0.5)
  #testx[2, 1:3] = c(-0.5, 0, -0.5)
  
  # get true survival function 
  timepoints = sort(unique(Y[Censor==1]))
  SurvMat = matrix(NA, nrow(testx), length(timepoints))
  exprate = xlink(testx)
  
  for (j in 1:length(timepoints))
  {
    SurvMat[, j] = 1 - pexp(timepoints[j], rate = exprate )
  }
```


```{r conf-band, fig.width=6, fig.height=6}
  RLTfit <- RLT(X, Y, Censor, model = "survival", 
                ntrees = 20000, nmin = 20, mtry = 20, split.gen = "random",
                resample.prob = 0.5, resample.replace = FALSE, nsplit = 5,
                param.control = list(split.rule = "logrank", "var.ready" = TRUE), 
                importance = FALSE, verbose = TRUE, ncores = 10)

  start_time <- Sys.time()
  RLTPred <- predict(RLTfit, testx, ncores = 10, var.est = TRUE)
  difftime(Sys.time(), start_time, units = "secs")
  
  paste("Prediction object size:", format(object.size(RLTPred), units = "MB"))
```

### Naive Monte Carlo

Red line is the truth, black-solid is estimated survival, black-dotted is +/- 1.96 of diagonal. blue-dotted is the naive Monte Carlo approach.

```{r class.source = NULL, fig.width=15, fig.height=25, out.width="90%"}
  alpha = 0.05

  # original Monte Carlo approach without smoothing
  start_time <- Sys.time()
  SurvBand = get.surv.band(RLTPred, alpha = alpha, approach = "eigen-th-mc")
  difftime(Sys.time(), start_time, units = "secs")

  par(mfrow = c(ceiling(ntest/3), 3))
  logt = log(1 + timepoints)
  
  for (i in 1:ntest)
  {
    # truth
    plot(logt, SurvMat[i, ], type = "l", lwd = 2, col = "red", ylab = paste("Subject",  i))
    
    # estimated and 1.96
    lines(logt, RLTPred$Survival[i,], lwd = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] - qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] + qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    
    # naive
    lines(logt, SurvBand[[i]]$lower[,1], lty = 2, lwd = 2, col = "deepskyblue")
    lines(logt, SurvBand[[i]]$upper[,1], lty = 2, lwd = 2, col = "deepskyblue")
  }
```

### Smoothed Rescaling MC Approach

```{r class.source = NULL, fig.width=15, fig.height=25, out.width="90%"}
  # Monte Carlo approach using smoothed rescaled cov matrix
  start_time <- Sys.time()
  SurvBand = get.surv.band(RLTPred, alpha = alpha, approach = "smoothed-mc", nsim = 1000)
  difftime(Sys.time(), start_time, units = "secs")

  par(mfrow = c(ceiling(ntest/3), 3))
  logt = log(1 + timepoints)
  
  for (i in 1:ntest)
  {
    # truth
    plot(logt, SurvMat[i, ], type = "l", lwd = 2, col = "red", ylab = paste("Subject",  i))
    
    # estimated and 1.96
    lines(logt, RLTPred$Survival[i,], lwd = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] - qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] + qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    
    # naive
    lines(logt, SurvBand[[i]]$lower[,1], lty = 2, lwd = 2, col = "deepskyblue")
    lines(logt, SurvBand[[i]]$upper[,1], lty = 2, lwd = 2, col = "deepskyblue")
  }
```

### Smoothed low-rank Monte Carlo

```{r class.source = NULL, fig.width=15, fig.height=25, out.width="90%"}
  # a low rank smoothed approach with adaptive bonferoni critical value
  start_time <- Sys.time()
  SurvBand = get.surv.band(RLTPred, alpha = alpha, approach = "smoothed-lr", r = 5)
  difftime(Sys.time(), start_time, units = "secs")

  par(mfrow = c(ceiling(ntest/3), 3))
  logt = log(1 + timepoints)
  
  for (i in 1:ntest)
  {
    # truth
    plot(logt, SurvMat[i, ], type = "l", lwd = 2, col = "red", ylab = paste("Subject",  i))
    
    # estimated and 1.96
    lines(logt, RLTPred$Survival[i,], lwd = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] - qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    lines(logt, RLTPred$Survival[i, ] + qnorm(1-alpha/2)*sqrt(diag(RLTPred$Cov[,,i])), 
          lty = 2, col = "black")
    
    # naive
    lines(logt, SurvBand[[i]]$lower[,1], lty = 2, lwd = 2, col = "deepskyblue")
    lines(logt, SurvBand[[i]]$upper[,1], lty = 2, lwd = 2, col = "deepskyblue")
  }
  
  cat(paste("average approximation error:", mean(sapply(SurvBand, "[[", 3))))
```

Here is a demonstration of the smoothed covariance approximation code used internally in the `get.surv.band` function. The following is the comparison of two covariance matrices

```{r fig.width=8, fig.height=8, out.width="45%"}
  i = 10
  ccov = RLTPred$Cov[,,i]
  heatmap(ccov[rev(1:nrow(ccov)),], Rowv = NA, Colv = NA, symm = TRUE)

  # generate grid points for kernel
  p = dim(ccov)[1]  
  alltime = 0:(p+1)
  alltime = alltime / stats::sd(alltime)
  nknots = ceiling( max(alltime)/orthoDr::silverman(1, p) ) + 2
  
  basis = orthoDr::kernel_weight(matrix(alltime), 
                                 matrix(seq(0, max(alltime), length.out = nknots)))
  
  # raw marginal variance
  mar_var = diag(ccov)
  
  fit <- glmnet::glmnet(basis, c(mar_var[1], mar_var, utils::tail(mar_var, 1)),
                        alpha = 0, intercept = FALSE,
                        lower.limits = 0, lambda = 1e-5)
  smarvar = stats::predict(fit, basis)
  smarvar = smarvar[2:(p+1)]
  
  newmat = ccov
  diag(newmat) = smarvar
  
  # find positive definite projection
  pdmat = suppressWarnings(Matrix::nearPD(newmat, 
                           keepDiag = TRUE, # keep the diagonal
                           base.matrix = TRUE, 
                           conv.norm.type = "F",
                           trace = FALSE, maxit = 200))
      
  sum((pdmat$mat - ccov)^2) / sum(ccov^2)
  heatmap(pdmat$mat[rev(1:nrow(pdmat$mat)),], Rowv = NA, Colv = NA, symm = TRUE)
```











