---
title: "RLT Package Testing Survival Functions and Features"
author: "Ruoqing Zhu"
date: "Last Updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true 
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r set-options, echo=FALSE, cache=FALSE}
  options(width = 1000)
  knitr::opts_chunk$set(fig.width=9, fig.height=7, out.width = "75%", fig.align = 'center')
  knitr::opts_chunk$set(class.source = "fold-show")
  knitr::opts_chunk$set(collapse=TRUE)
```

## Install and Load Package

Install and load the GitHub version of the RLT package. Do not use the CRAN version. 

```{r}
  # install.packages("devtools")
  # devtools::install_github("teazrq/RLT")
  library(RLT)
```

Load other packages used in this guide.

```{r setup, include=FALSE}
  pkgs <- c("randomForest", "randomForestSRC", "ranger", "parallel")
  for (p in pkgs) if (!requireNamespace(p, quietly = TRUE)) install.packages(p)
```
```{r message=FALSE}  
  library(randomForest)
  library(randomForestSRC)
  library(ranger)
  library(parallel)
```

## Benchmark Against Existing Packages

We generate a dataset with 300 observations and 200 variables, with 100 continuous variables and 100 categorical ones with three categories. The survival time and censoring time all follow an exponential distribution. 

```{r generate-data}
  # generate date
  set.seed(1)
  trainn = 500
  testn = 1000
  n = trainn + testn
  p = 200
  X1 = matrix(rnorm(n*p/2), n, p/2)
  X2 = matrix(as.integer(runif(n*p/2)*3), n, p/2)

  X = data.frame(X1, X2)
  xlink <- function(x) exp(x[, 7] + x[, 16] + x[, 25] + x[, p])
  FT = rexp(n, rate = xlink(X))
  CT = rexp(n, rate = 0.5)
  
  y = pmin(FT, CT)
  Censor = as.numeric(FT <= CT)
  mean(Censor)
  
  # parameters
  ntrees = 500
  ncores = 10
  nmin = 25
  mtry = p/3
  sampleprob = 0.85
  rule = "random"
  nsplit = ifelse(rule == "best", 0, 3)
  importance = TRUE
  
  trainX = X[1:trainn, ]
  trainY = y[1:trainn]
  trainCensor = Censor[1:trainn]
  
  testX = X[1:testn + trainn, ]
  testY = y[1:testn + trainn]
  testCensor = Censor[1:testn + trainn]
  
  # get true survival function 
  timepoints = sort(unique(trainY[trainCensor==1]))
  
  SurvMat = matrix(NA, testn, length(timepoints))
  expxlink = xlink(testX) 
    
  for (j in 1:length(timepoints))
  {
    SurvMat[, j] = 1 - pexp(timepoints[j], rate = expxlink )
  }
  
  # Calculate timepoint indices for ranger (different from RLT)
  yloc = rep(NA, length(timepoints))
  for (i in 1:length(timepoints)) yloc[i] = sum( timepoints[i] >= trainY )
  
  for (j in (p/2 + 1):p) X[,j] = as.factor(X[,j])
```


```{r speed-comparison, class.source = NULL}
  # recording results
  metric = data.frame(matrix(NA, 6, 7))
  rownames(metric) = c("rlt", "rltsup", "rltcox", "rltcoxpen", "rsf", "ranger")
  colnames(metric) = c("fit.time", "pred.time", "oob.error", "pred.error", "L1", 
                       "obj.size", "tree.size")
  
  # fit RLT with log-rank split
  start_time <- Sys.time()
  
  RLTfit.logrank <- RLT(trainX, trainY, trainCensor, model = "survival", 
                        ntrees = ntrees, ncores = ncores, 
                        nmin = nmin, mtry = mtry, nsplit = nsplit,
                        split.gen = rule, resample.prob = sampleprob,
                        importance = importance, 
                        param.control = list(split.rule = "logrank", "alpha" = 0.2), 
                        verbose = TRUE, resample.replace=FALSE)
  metric[1, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  RLTPred <- predict(RLTfit.logrank, testX, ncores = ncores)
  metric[1, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[1, 3] = RLTfit.logrank$Error
  metric[1, 4] = 1 - cindex(testY, testCensor, colSums(apply(RLTPred$Hazard, 1, cumsum)))
  metric[1, 5] = mean(colMeans(abs(RLTPred$Survival - SurvMat)))
  metric[1, 6] = format(object.size(RLTfit.logrank), units = "MB")
  metric[1, 7] = mean(unlist(lapply(RLTfit.logrank$FittedForest$SplitVar, length)))
  
  # fit RLT with sup-log-rank split  
  start_time <- Sys.time()
  RLTfit.suplogrank <- RLT(trainX, trainY, trainCensor, model = "survival", 
                           ntrees = ntrees, ncores = ncores, 
                           nmin = nmin, mtry = mtry, nsplit = nsplit,
                           split.gen = rule, resample.prob = sampleprob,
                           importance = importance, 
                           param.control = list(split.rule = "suplogrank", "alpha" = 0), 
                           verbose = TRUE, resample.replace=FALSE)
  metric[2, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  RLTPred <- predict(RLTfit.suplogrank, testX, ncores = ncores)
  metric[2, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[2, 3] = RLTfit.suplogrank$Error
  metric[2, 4] = 1- cindex(testY, testCensor, colSums(apply(RLTPred$Hazard, 1, cumsum)))
  metric[2, 5] = mean(colMeans(abs(RLTPred$Survival - SurvMat)))
  metric[2, 6] = format(object.size(RLTfit.suplogrank), units = "MB")
  metric[2, 7] = mean(unlist(lapply(RLTfit.suplogrank$FittedForest$SplitVar, length)))
  
  # fit RLT with cox-grad split  
  start_time <- Sys.time()
  RLTfit.cg <- RLT(trainX, trainY, trainCensor, model = "survival", 
                   ntrees = ntrees, ncores = ncores, 
                   nmin = nmin, mtry = mtry, nsplit = nsplit,
                   split.gen = rule, resample.prob = sampleprob,
                   importance = importance,
                   param.control = list(split.rule = "coxgrad", "alpha" = 0), 
                   verbose = TRUE, resample.replace=FALSE)
  metric[3, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  RLTPred <- predict(RLTfit.cg, testX, ncores = ncores)
  metric[3, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[3, 3] = RLTfit.cg$Error
  metric[3, 4] = 1 - cindex(testY, testCensor, colSums(apply(RLTPred$Hazard, 1, cumsum)))
  metric[3, 5] = mean(colMeans(abs(RLTPred$Survival - SurvMat)))
  metric[3, 6] = format(object.size(RLTfit.cg), units = "MB")
  metric[3, 7] = mean(unlist(lapply(RLTfit.cg$FittedForest$SplitVar, length)))
  
  # fit RLT with penalized coxgrad split
  start_time <- Sys.time()
  RLTfit.pcg <- RLT(trainX, trainY, trainCensor, model = "survival", 
                    ntrees = ntrees, nmin = nmin, mtry = mtry, nsplit = nsplit,
                    split.gen = rule, resample.prob = sampleprob, importance = importance, 
                    # var.w = ifelse(c(1:(p)) %in% c(7, 16, 25, p), 1, 0.5),
                    var.w = pmax(max(0, mean(RLTfit.logrank$VarImp)), RLTfit.logrank$VarImp),
                    param.control = list(split.rule = "coxgrad", "alpha" = 0), 
                    verbose = TRUE, ncores = ncores, resample.replace=FALSE)
  metric[4, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  RLTPredp <- predict(RLTfit.pcg, testX, ncores = ncores)
  metric[4, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[4, 3] = RLTfit.pcg$Error
  metric[4, 4] = 1 - cindex(testY, testCensor, colSums(apply(RLTPredp$Hazard, 1, cumsum)))
  metric[4, 5] = mean(colMeans(abs(RLTPredp$Survival - SurvMat)))
  metric[4, 6] = format(object.size(RLTfit.pcg), units = "MB")
  metric[4, 7] = mean(unlist(lapply(RLTfit.pcg$FittedForest$SplitVar, length)))
  
  # fit rsf
  options(rf.cores = ncores)
  start_time <- Sys.time()
  rsffit <- rfsrc(Surv(trainY, trainCensor) ~ ., data = data.frame(trainX, trainY, trainCensor),
                  ntree = ntrees, nodesize = nmin, mtry = mtry,
                  nsplit = nsplit, sampsize = trainn*sampleprob, 
                  importance = ifelse(importance==TRUE,"random", "none"), samptype = "swor",
                  block.size = 1, ntime = NULL)
  metric[5, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  rsfpred = predict(rsffit, data.frame(testX))
  metric[5, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[5, 3] = tail(rsffit$err.rate, 1)
  metric[5, 4] = 1 - cindex(testY, testCensor, rowSums(rsfpred$chf))
  metric[5, 5] = mean(colMeans(abs(rsfpred$survival - SurvMat)))
  metric[5, 6] = format(object.size(rsffit), units = "MB")
  metric[5, 7] = rsffit$forest$totalNodeCount / rsffit$forest$ntree
  
  # fit ranger
  start_time <- Sys.time()
  rangerfit <- ranger(Surv(trainY, trainCensor) ~ ., data = data.frame(trainX, trainY, trainCensor), 
                      num.trees = ntrees, min.node.size = nmin, mtry = mtry, 
                      splitrule = "logrank", num.threads = ncores, 
                      sample.fraction = sampleprob, importance = "permutation")
  metric[6, 1] = difftime(Sys.time(), start_time, units = "secs")
  start_time <- Sys.time()
  rangerpred = predict(rangerfit, data.frame(testX))
  metric[6, 2] = difftime(Sys.time(), start_time, units = "secs")
  metric[6, 3] = rangerfit$prediction.error
  metric[6, 4] = 1 - cindex(testY, testCensor, rowSums(rangerpred$chf))
  # For ranger, calculate L1 error using available timepoints
  # Use the first min(ncol(rangerpred$survival), ncol(SurvMat)) columns
  n_cols = min(ncol(rangerpred$survival), ncol(SurvMat))
  metric[6, 5] = mean(colMeans(abs(rangerpred$survival[, 1:n_cols] - SurvMat[, 1:n_cols])))
  metric[6, 6] = format(object.size(rangerfit), units = "MB")
  metric[6, 7] = mean(unlist(lapply(rangerfit$forest$split.varIDs, length)))
  
  metric
```



## Print a Single Tree

You can use the `get.one.tree()` function to peek into a single tree.  

```{r print-tree}
  # print one tree
  tree_num = 1
  get.one.tree(RLTfit.logrank, tree_num)

  # to get the estimated hazard function of a terminal node
  terminal_nodes = RLTfit.logrank$FittedForest$SplitVar[[tree_num]] == -1
  haz = RLTfit.logrank$FittedForest$NodeHaz[[tree_num]][[which(terminal_nodes)[1]]]
  
  # to get the estimated survival function of a terminal node
  plot(c(0, RLTfit.logrank$timepoints),
       exp(-cumsum(haz)),
       xlab = "time", ylab = "survival", type = "l", 
       main = paste("First terminal node of Tree", tree_num))
```

## Random Forest Kernel

```{r kernel}
  KernelW = forest.kernel(RLTfit.logrank, X1 = testX[1, ], X2 = trainX)$Kernel
```

## Setting random seed

```{r}
  ## Fitting a forest
  RLTfit1 <- RLT(trainX, trainY, trainCensor, model = "survival", 
                 ntrees = 100, importance = TRUE, nmin = 1)

  RLTfit2 <- RLT(trainX, trainY, trainCensor, model = "survival", 
                 ntrees = 100, importance = TRUE, nmin = 1,
                 seed = RLTfit1$parameters$seed)
  
  # check if importance are identical
  all(RLTfit1$VarImp == RLTfit2$VarImp)
  
  # prediction
  RLTPred1 <- predict(RLTfit1, testX, keep.all = TRUE)
  RLTPred2 <- predict(RLTfit2, testX, keep.all = TRUE)

  # check predictions are identical
  all(RLTPred1$Prediction == RLTPred2$Prediction)
```

## Categorical Variables 

This section demonstrates how RLT handles categorical variables in survival analysis.