---
title: "Survival — Minimal Tutorial (RLT)"
author: "Ruoqing Zhu"
date: "Last Updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true 
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
# Clean, consistent output for all chunks
knitr::opts_chunk$set(
  message = FALSE, warning = FALSE,
  fig.width = 7, fig.height = 5, out.width = "70%", fig.align = "center",
  collapse = TRUE, class.source = "fold-show"
)
```

## Overview

This page shows how to fit and predict a survival model with RLT.

## Prerequisites

See [Get Started](articles/Test-RLT.html) for installation and loading instructions.

## Data

We generate continuous and integer predictors with right-censored survival times. The dataset includes event times following an exponential model with rate depending on predictors, and censoring times that control the censoring level. We observe time = min(event_time, censor_time) and status = 1(event before censor).

```{r data-prep}
# (Optional) For reproducibility in this tutorial only.
# Detailed notes on random seeds are in the Seed & Reproducibility feature page.
set.seed(1)

# ---- Generate a small synthetic dataset ----
trainn <- 80
testn  <- 20
n <- trainn + testn
p <- 10

# Continuous + integer "categorical" predictors
X1 <- matrix(rnorm(n * (p/2)), n, p/2)
X2 <- matrix(as.integer(runif(n * (p/2)) * 3), n, p/2)  # integers 0,1,2

X <- data.frame(X1, X2)

# Link for event rate (uses a few columns to create signal)
xlink <- function(x) exp(0.6 * x[, 2] + 0.6 * x[, 4] + 0.4 * x[, p])

# Event and censoring times
FT <- rexp(n, rate = xlink(X))   # event (failure) time
CT <- rexp(n, rate = 0.5)        # censoring time (controls censoring level)

# Observed time and status (1 = event observed, 0 = censored)
time   <- pmin(FT, CT)
status <- as.numeric(FT <= CT)

# Train / test split
trainX     <- X[1:trainn, ]
trainTime  <- time[1:trainn]
trainStat  <- status[1:trainn]
testX      <- X[(trainn + 1):(trainn + testn), ]
testTime   <- time[(trainn + 1):(trainn + testn)]
testStat   <- status[(trainn + 1):(trainn + testn)]
```

## Fit

```{r fit-model}
ntrees <- 200
ncores <- max(1, parallel::detectCores() - 1)
nmin   <- 5
mtry   <- p/2
samplereplace <- TRUE
sampleprob    <- 0.80
rule    <- "best"     # split generator
nsplit  <- 0
importance <- TRUE

fit <- RLT(
  trainX, trainTime, trainStat, model = "survival",
  ntrees = ntrees, mtry = mtry, nmin = nmin,
  resample.prob = sampleprob, split.gen = rule,
  resample.replace = samplereplace,
  nsplit = nsplit, importance = importance,
  param.control = list(alpha = 0),
  ncores = ncores, verbose = TRUE
)
```

## Predict

Predict survival and hazard estimates for the test set; compute a simple risk score for evaluation.

```{r predict}
pred <- predict(fit, testX, keep.all = TRUE, ncores = ncores)

# Optional helpers depending on returned structure
# Risk score: sum of cumulative hazards (a simple concordance-friendly score)
risk_score <- colSums(apply(pred$Hazard, 1, cumsum))
```

## Evaluate

```{r evaluate}
c_index <- 1 - cindex(testTime, testStat, risk_score)

# A compact summary
list(
  Concordance = round(c_index, 4)
)
```

## Inspect

```{r inspect}
print(fit)
```