---
title: "RLT Package Reinforcement Learning Test"
author: "Ruoqing Zhu"
date: "Last Updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: yes
    toc_float:
      collapsed: true
      smooth_scroll: true
    toc_depth: '2'
  pdf_document:
    toc: yes
    toc_depth: 2
---

```{r set-options, echo=FALSE, cache=FALSE}
  options(width = 1000)
  knitr::opts_chunk$set(fig.width=9, fig.height=7, out.width = "75%", fig.align = 'center')
  knitr::opts_chunk$set(class.source = "fold-show")
  knitr::opts_chunk$set(collapse=TRUE)
```

## Install the RLT Package

### For Windows Users

In Windows, RLT can be installed using `install_github`. The compilation of this package requires GCC. Hence, you may also need to install Rtools from [here](https://cran.r-project.org/bin/windows/Rtools/).

```{r, eval=FALSE}
  # you may need to install the devtools package first
  # install.packages("devtools")
  library(devtools)

  # install the RLT package from GitHub
  # Rtools is required in Windows
   install_github("teazrq/RLT")
```

### For macOS Users

When installing RLT on macOS, the process can be slightly trickier. It mainly consists of two steps:

#### Step 1: Install Compilers

- Follow [this guide](https://mac.r-project.org/tools/) to install the GNU Fortran compiler: [gfortran-12.2-universal.pkg](https://mac.r-project.org/tools/gfortran-12.2-universal.pkg)
- If you do not already have Xcode (most systems come with Xcode installed already), you can use this line to install it:

```sh
sudo xcode-select --install
```

#### Step 2: Set Makevars to Point to the Compiler

Create the Makevars file using:

```sh
mkdir ~/.R
touch ~/.R/Makevars
```

Check your folder of gfortran. The compiler will be installed to `/opt/gfortran` folder as default.

Add these lines to point to the correct folder of the compiler. You may use `open -a TextEdit ~/.R/Makevars` to open a text editor and type in these lines:

```
FC = /opt/gfortran/bin/gfortran
F77 = /opt/gfortran/bin/gfortran
FLIBS = -L/opt/gfortran/lib
```

After completing these steps, you should be able to directly install the RLT package using `install_github("teazrq/RLT")` in RStudio the same way as Windows machine. The compilation may require a few minutes.

### OpenMP in macOS

The previous steps would not activate OpenMP parallel computing. To enable OpenMP while compiling the RLT package, follow these steps:

#### Step 1: Download and Install LLVM

For example, to install the 14.0.6 version (check the ones that compatible with your Xcode version), use:

```sh
curl -O https://mac.r-project.org/openmp/openmp-14.0.6-darwin20-Release.tar.gz
sudo tar fvxz openmp-14.0.6-darwin20-Release.tar.gz -C /
```

You should then see the following message:

```
usr/local/lib/libomp.dylib
usr/local/include/ompt.h
usr/local/include/omp.h
usr/local/include/omp-tools.h
```

#### Step 2: Add Flags into Makevars

The procedure is the same as we explained previously. You should add these two lines into your Makevars file:

```
CPPFLAGS += -Xclang -fopenmp
LDFLAGS += -lomp
```

After these two steps, you can then use `install_github("teazrq/RLT", force = TRUE)` to re-install the RLT package.

### Load the RLT Package

```{r}
  library(RLT)
```

## Benchmarking

The following code tests the performance of the RLT package in a regression problem:

```{r}
  library(parallel)
  # Set seed for reproducibility
  set.seed(1)

  # Define data size
  trainn <- 800
  testn <- 1000
  n <- trainn + testn
  p <- 30

  # Generate continuous variables (X1) and categorical variables (X2)
  X1 <- matrix(rnorm(n * p / 2), n, p / 2)
  X2 <- matrix(as.integer(runif(n * p / 2) * 3), n, p / 2)

  # Combine continuous and categorical variables into a data frame (X)
  X <- data.frame(X1, X2)

  # Convert the second half of the columns in X to factors
  X[, (p / 2 + 1):p] <- lapply(X[, (p / 2 + 1):p], as.factor)

  # Generate outcomes (y)
  y <- 1 + rowSums(X[, 2:6]) + 2 * (X[, p / 2 + 1] %in% c(1, 3)) + rnorm(n)

  # Set tuning parameters
  ntrees <- 1000
  ncores <- detectCores() - 1
```

```{r}
  # Benchmark RLT
  start_time <- Sys.time()
  
  RLTfit <- RLT(X[1:trainn, ], y[1:trainn], model = "regression",
                ntrees = ntrees, ncores = ncores, nmin = 10,
                split.gen = "random", nsplit = 1,
                resample.prob = 0.85, resample.replace = FALSE,
                reinforcement = TRUE, importance = TRUE,
                param.control = list("embed.ntrees" = 50,
                                     "embed.mtry" = 1/2,
                                     "embed.nmin" = 10,
                                     "embed.split.gen" = "random",
                                     "embed.nsplit" = 1,
                                     "embed.resample.replace" = TRUE,
                                     "embed.resample.prob" = 0.9,
                                     "embed.mute" = 0.5,
                                     "embed.protect" = 5),
                verbose = TRUE)
  
  RLT_time <- difftime(Sys.time(), start_time, units = "secs")
  
  # Make predictions
  RLT_pred <- predict(RLTfit, X[(trainn + 1):n, ])
  RLT_error <- mean((RLT_pred$Prediction - y[(trainn + 1):n])^2)
  
  cat("RLT Results:\n")
  cat("Training time:", round(as.numeric(RLT_time), 2), "seconds\n")
  cat("Test MSE:", round(RLT_error, 4), "\n")
  cat("Variable importance (top 5):\n")
  print(head(sort(RLTfit$VarImp, decreasing = TRUE), 5))
```

```{r fig.width=8, fig.height=6}
  # Plot variable importance
  par(mar = c(5, 8, 4, 2))
  barplot(sort(RLTfit$VarImp, decreasing = TRUE)[1:10], 
          horiz = TRUE, las = 1, 
          main = "RLT Variable Importance (Top 10)",
          xlab = "Importance Score")
```

## Single Variable Embedded Splitting

```{r}
 # set.seed(2)
  
  n = 1000
  p = 10
  X = matrix(rnorm(n*p), n, p)
  y = 1 + X[, 1] + X[, 3] + X[, 9] + rnorm(n)
  
  testX = matrix(rnorm(n*p), n, p)
  testy = 1 + testX[, 1] + testX[, 3] + testX[, 9] + rnorm(n)
  
  start_time <- Sys.time()
  
  RLTfit <- RLT(X, y, model = "regression", 
                ntrees = 100, ncores = 1, nmin = 10,
                split.gen = "random", nsplit = 1,
                resample.prob = 0.85, resample.replace = FALSE,
                reinforcement = TRUE, importance = "distribute",
                param.control = list("embed.ntrees" = 50,
                                     "embed.mtry" = 1/2,
                                     "embed.nmin" = 10,
                                     "embed.split.gen" = "random",
                                     "embed.nsplit" = 1,
                                     "embed.resample.replace" = TRUE,
                                     "embed.resample.prob" = 0.9,
                                     "embed.mute" = 0.5,
                                     "embed.protect" = 5),
                verbose = TRUE)
  
  difftime(Sys.time(), start_time, units = "secs")
  
    # oob error
  mean((RLTfit$Prediction - y)^2, na.rm = TRUE)
  
  # prediction error
  pred = predict(RLTfit, testX)
  mean((pred$Prediction - testy)^2)
  
```

```{r fig.width= 8, fig.height=5}
  # sparse variable importance
  barplot(as.vector(RLTfit$VarImp), main = "RLT")
  
  # check one tree
  get.one.tree(RLTfit, 1)
```

Check seed match

```{r}
  RLTfit2 <- RLT(X, y, model = "regression", 
                 ntrees = 20, ncores = 1, nmin = 10,
                 split.gen = "random", nsplit = 1,
                 resample.prob = 0.85, resample.replace = FALSE,
                 reinforcement = TRUE, importance = TRUE,
                 param.control = list("embed.ntrees" = 50,
                                      "embed.mtry" = 1/3,
                                      "embed.nmin" = 10,
                                      "embed.split.gen" = "random",
                                      "embed.nsplit" = 1,
                                      "embed.resample.prob" = 0.75,
                                      "embed.mute" = 0.5,
                                      "embed.protect" = 5),
                 verbose = TRUE, seed = RLTfit$parameters$seed)

  all(RLTfit$VarImp == RLTfit2$VarImp)
```

## Linear Combination Split 

This is a model without fitting embedded random forest to search for the linear combinations. The variables are ranked by marginal screening. 

```{r}
  # set.seed(1)
  library(MASS)
  n = 1000
  p = 5
  S = matrix(0.3, p, p)
  diag(S) = 1
  S[1, 5] = S[5, 1] = 0.9
  S[1, 3] = S[3, 1] = S[5, 3] = S[3, 5] = -0.3
  
  X1 = mvrnorm(n, mu = rep(0, p), Sigma = S)
  X2 = as.factor(sample(c(-1,1), size = n, replace = TRUE))
  X = data.frame(X1, X2)
  y = 1 + 2*(X[, 6] == 1) + X[, 1] + X[, 3] + rnorm(n)
  w = runif(n)
  
  start_time <- Sys.time()
  
  RLTfit <- RLT(X, y, model = "regression", obs.w = w,
                ntrees = 100, ncores = 1, nmin = 50, mtry = 3,
                split.gen = "random", nsplit = 3, 
                resample.prob = 0.8, resample.replace = FALSE, 
                split.rule = "sir",
                param.control = list("linear.comb" = 3), #ZRL
                importance = TRUE, 
                verbose = TRUE)
  
  difftime(Sys.time(), start_time, units = "secs")
  
  plot(RLTfit$Prediction, y)
  mean( (RLTfit$Prediction - y)^2 , na.rm = TRUE)
```